<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Hybrid Evaluation}
-->

```{r}
knitr::opts_chunk$set(tidy = FALSE)
library(predcomps)
library(ggplot2)
library(reshape2)
library(plyr)
```

This example will have 9 inputs: $v$ and $u_1$, $u_2$, ..., $u_8$. Each $u$ can transition from one value to another only at one value of $v$.  Each $u$ will have the same role in the model, which is $$y ~ vu_1 + vu_2 + vu_3 + vu_4 + vu_5 + vu_6 + vu_7 + vu_8)$$.

I'll construct the data and show the relationship between the $u$'s and $v$:
```

```{r}
N <- 100
vValues <- (-3):3
v <- sample(vValues, N, replace=TRUE)


df <- data.frame(v)
for (i in seq_along(vValues)) {
  df[[paste0("u",i)]] <- ifelse(v==vValues[i], 
                                sample(c(0,10), N, replace=TRUE),
                                rep(0, N)
                                )
}
df$u8 <- ifelse(v %in% c(-3,3),
                sample(c(0,10), N, replace=TRUE),
                rep(0, N)
                )

ggplot(melt(df, id="v")) +
  geom_bar(aes(x=factor(v), fill=factor(value)), position=position_fill()) +
  facet_grid(variable ~ .) +
  scale_fill_discrete("u_i value") + 
  scale_x_discrete("v") +
  ggtitle("Distributions of each u conditional on v")
```

You can see that the $u$'s are almost always constant, with transitions happening at only one value of $v$ for each $u$, except that $u_8$ can transition at either $v=-3$ or $v=3$.

```{r}
outputGenerationFunction <- function(df) {
  with(df, v*u1 + v*u2 + v*u3 + v*u4 + v*u5 + v*u6 + v*u7 + v*u8)
}
df$y <- outputGenerationFunction(df)
```

Compute the APCs:

```{r}
inputVars <- c("v",paste0("u",1:8))
apcDF <- GetApcDF(outputGenerationFunction, df, inputVars)

longAPCs <- melt(apcDF, id="Input", value.name = "APC", variable.name = "Type")


maxAPC <- max(abs(longAPCs$APC))

longAPCs2 <- rbind(
  longAPCs,
  transform(subset(longAPCs, Type=="Absolute"), APC=-APC)
)
longAPCs2$Input <- reorder(factor(longAPCs2$Input), longAPCs2$APC, FUN = function(x) mean(abs(x))) 

longAPCs2 <- longAPCs2[order(factor(longAPCs2$Type, levels=c("Absolute", "Signed"))), ]

ggplot(longAPCs2) +
  geom_point(aes(y = Input, x=APC, color=Type, shape=Type, size=Type)) +
  #facet_grid(. ~ Type) + 
  theme_grey(base_size=20) +
  scale_x_continuous(limits=c(-maxAPC, maxAPC)) +
  scale_size_discrete(range=c(3,4))

## I like this way of plotting APCs!
```


